Increased Education Level of Women Associates with Higher Occupational Prestige and Happiness

### 2023-03-16 outputs/paper/paper.qmd
final edits

According to the World Bank [@citeWorldBank], the education of girls enhances economies and mitigates inequality, promoting more stable and resilient societies that offer all individuals, including boys and men, the chance to reach their full potential. Over the past two decades, there has been a steady increase in women's education levels, as evidenced by Figure 1, which displays a positive upward trajectory in women's education levels over time. This trend can be attributed to several factors.

@@

The results from @fig-2 shows a positive trend in the level of prestige of women\'s occupations had a generally upward trend, meaning that women\'s occupation prestige is increasing over time. There are several factors that contribute to occupational prestige scores for women. Higher levels of education, such as obtaining a bachelor's or graduate degree, are generally associated with higher career prestige scores. Women with higher income levels may also have higher prestige scores. Certain job titles, such as CEO or doctor, may be associated with higher levels of prestige for women. Women who work in certain professions, such as lawyers or professors, may be considered more prestigious due to their perceived social status. Additionally, certain industries, such as finance or law, may be considered more prestigious than others. Historical and cultural factors may also influence the perception of occupational prestige for women, with certain occupations, such as being a doctor or lawyer, considered more prestigious than others.

@@

Our results from @fig-2 indicate that the mean occupational prestige score increased by over 3% suggesting that there are more women working and more work looks like work that would have been associated with women in the past. According to citation, the researchers at the University of Chicago and the Chicago Federal Reserve found that over time, women have become happier and more satisfied with their work, while men have become less happy and less satisfied. The researchers offer several reasons for this gender difference. First, changing social norms have allowed more women to enter the workplace in recent decades. Second, the work options available for women have expanded dramatically since the 1950s, with more professional and managerial jobs becoming available. Third, more women than men now graduate in the United States with college and graduate school degrees, which increases their options. Finally, lower-educated women have experienced the greatest increase in workplace satisfaction, possibly because they were the most constrained under old gendered rules.

@@

@fig-1, @fig-2 and @fig-3 show growth in their values, with @fig-3 specifically showing growth with the values correlated to being happy. We can reasonably deduce that there is some positive relationship between happiness, the prestige of a woman's occupation and a woman\'s peak education level. From [@citeLinked], the work options for women have expanded due to an increase in education and societal factors. An increase in work options means that women are shifting into more professional jobs rather than clerical ones. These factors lead to an increase in women's happiness overall.

@@

There is selection bias with the women that willingly participate in the survey. We are using the responses of randomly selected women that voluntarily fill out the women to make blanket statements about all women in America. Women that fill out the GSS could differ from the majority of women in America, but because they are the ones that filled out the GSS, we are using their data to generalize the whole populace. Naturally, using data from a subset of our estimand to analyze the whole estimand will produce biased results.

Measurement bias is also a factor here. The GSS survey is conducted face-to-face with an in-person interview. This means that some of the results found in the GSS could\'ve been influenced by having a face-to-face interview.. Some of the questions asked in the GSS may have different answers if they were asked virtually, or through the web, instead.

@@

Our paper analyzes the highest education level of women in relation to their occupational prestige and happiness. However, our data only examines the education level of women in the US and is insufficient to support the analysis of women\'s general level of higher education in the world. To further strengthen our findings, we can take the GSS of other Western countries, such as Canada, and compare the data for a more multifaceted outlook. Moreover, according to Appendix G of the GSS: Prestige Score and Socioeconomic Index Distributions, the rating systems for the prestige scores were first developed through a study in 19063-1965, then updated in 1989. As society\'s view on the social standing of occupations has changed significantly since 1989, the old metrics used for our analysis may not accurately reflect the standards of prestige in the modern day.

>>>>

According to the World Bank [@citeWorldBank], the education of girls enhances economies and mitigates inequality, promoting more stable and resilient societies that offer all individuals, including boys and men, the chance to reach their full potential. Over the past two decades, there has been a steady increase in women's education levels, as evidenced by @fig-1, which displays a positive upward trajectory in women's education levels over time. This trend can be attributed to several factors.

@@

The results from @fig-2 shows a positive trend in the level of prestige of women\'s occupations had a generally upward trend, meaning that women\'s occupation prestige is increasing over time. There are several factors that contribute to occupational prestige scores for women. Higher levels of education, such as obtaining a Bachelor's or Graduate degree, are generally associated with higher career prestige scores. Women with higher income levels may also have higher prestige scores. Certain job titles, such as CEO or doctor, may be associated with higher levels of prestige for women. Women who work in certain professions, such as lawyers or professors, may be considered more prestigious due to their perceived social status. Additionally, certain industries, such as finance or law, may be considered more prestigious than others. Historical and cultural factors may also influence the perception of occupational prestige for women, with certain occupations, such as being a doctor or lawyer, considered more prestigious than others.

@@

Our results from @fig-2 indicate that the mean occupational prestige score increased by over 3% suggesting that there are more women working and more work looks like work that would have been associated with women in the past. According to Litwin, researchers at the University of Chicago and the Chicago Federal Reserve found that over time, women have become happier and more satisfied with their work, while men have become less happy and less satisfied [@citeLinked]. The researchers offer several reasons for this gender difference. First, changing social norms have allowed more women to enter the workplace in recent decades. Second, the work options available for women have expanded dramatically since the 1950s, with more professional and managerial jobs becoming available. Third, more women than men now graduate in the United States with college and graduate school degrees, which increases their options. Finally, lower-educated women have experienced the greatest increase in workplace satisfaction, possibly because they were the most constrained under old gendered rules.

@@

@fig-1, @fig-2, and @fig-3 all show growth in their values, with @fig-3 specifically showing growth with the values correlated to happiness. We can reasonably deduce that there is some positive relationship between happiness, the prestige of a woman's occupation and a woman\'s peak education level. From [@citeLinked], the work options for women have expanded due to an increase in education and societal factors. An increase in work options means that women are shifting into more professional jobs rather than clerical ones. These factors lead to an increase in women's happiness overall.

@@

There is selection bias with the women who willingly participate in the survey. We are using the responses of randomly selected women who voluntarily fill out the survey to make blanket statements about all women in America. Women that fill out the GSS could differ from the majority of women in America, but because they are the ones that filled out the GSS, we are using their data to generalize the whole populace. Naturally, using data from a subset of our estimand to analyze the whole estimand will produce biased results.

Measurement bias is also a factor here. The GSS survey is conducted face-to-face with an in-person interview. This means that some of the results found in the GSS could\'ve been influenced by having a face-to-face interview. Some of the questions asked in the GSS may have different answers if they were asked virtually, or through the web, instead.

@@

Our paper analyzes the highest education level of women in relation to their occupational prestige and happiness. However, our data only examines the education level of women in the US and is insufficient to support the analysis of women\'s general level of higher education in the world. To further strengthen our findings, we can take the GSS of other Western countries, such as Canada, and compare the data for a more multifaceted outlook. Moreover, according to Appendix G of the GSS: Prestige Score and Socioeconomic Index Distributions, the rating systems for the prestige scores were first developed through a study in 19063-1965, then updated in 1989 [@citeGSS]. As society\'s view on the social standing of occupations has changed significantly since 1989, the old metrics used for our analysis may not accurately reflect the standards of prestige in the modern day.
